\section{Theory}

Our goal is to find the ground state energy of a helium atom, using the variational Monte Carlo method. To this end we will use trial wave functions $\psi_T(\textbf{r},\alpha)$, which is dependent on position $\textbf{r}$ and some parameter $\alpha$. Notice that it is not dependent on time; we are merely interested in the ground state energy and that does not evolve over time. In order to find the expectation value of the ground state energy we can use equation \ref{eq_groundstate}: 

\begin{equation}\label{eq_groundstate}
E(\alpha) = \frac{\int d\textbf{r} \psi_T^*(\textbf{r},\alpha)H\psi_T(\textbf{r},\alpha)}{\int d\textbf{r} \psi_T^*(\textbf{r},\alpha)\psi_T(\textbf{r},\alpha)}
\end{equation}

Here $H$ is the Hamiltonian and $E(\alpha) \geq E_0$ is the approximation of the ground state energy which is always higher than or equal to the actual ground state energy $E_0$.  We will also define the \textit{local energy} $E_L = \frac{H\psi_T(\textbf{r},\alpha)}{\psi_T(\textbf{r},\alpha)}$ which is a flat function in case the $\psi_T(\textbf{r},\alpha) = \psi_0(\textbf{r})$. By evaluating the variance of $E_L$, we can approximate how far from the ground state energy we are. If the variance is equal to zero we have found an exact solution. Meanwhile we can make use this local energy as well, by simplifying equation \ref{eq_groundstate} to equation \ref{eq_groundstate2}

\begin{equation}\label{eq_groundstate2}
E(\alpha) = \frac{\int d\textbf{r} \psi_T^*(\textbf{r},\alpha)\psi_T(\textbf{r},\alpha)E_L(\textbf{r},\alpha)}{\int d\textbf{r} \psi_T^*(\textbf{r},\alpha)\psi_T(\textbf{r},\alpha)}
\end{equation}

which is easier to work with due to the lack of a Hamiltonian. [REFERENCE TO BOOK] These integrals are then evaluated using Monte Carlo simulation. \\

In order to make the best use of finite integration steps, we will sample not \textit{entirely} at random, but rather by iteratively in such a way that we place our finite sample points there where the wave function is non-zero. This is called an adaptive method, and we will show that this yields better results quicker than using an entirely random sample. For further details see chapter 3. In order to calculate the variance of the local energy, we take $Var(E_L) = \sqrt{<E_L^2> - <E_L>^2} = \sqrt{ \int \textbf{r}^2 E_L(\textit{r},\alpha) d\textbf{r} - (\int \textbf{r} E_L(\textit{r},\alpha)d\textbf{r})^2 } $, which can either be solves analytically or numerically, depending on the form of the local energy. 